Understanding Parameterized Quantum Circuit Learning for Quantum Chemical Applications

### 2025-04-07 manuscript/main.tex
Fri  4 Apr 2025 14:29:29 EDT
Mon  7 Apr 2025 15:56:12 EDT

The last experiments we use state-vector for are learning curves since previous studies have mentioned that PQCs can outperform traditional models with fewer data points \cite{suzuki_predicting_2020,hatakeyama-sato_quantum_2023} \textbf{[double check]}
\siref{table:5BSE49_LC}
\siref{table:DDCC_LC}

Error bars show the min and max values

\begin{figure}[H]
	\centering	
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/BSE/fivequbit/BSE_learningcurve}
		\caption{}
		\label{fig:BSE5_learning_curves}
	\end{subfigure}	
	\hfill
	\begin{subfigure}{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/DDCC/DDCC_learning_curves}
		\caption{}
		\label{fig:ddcclearningcurves}
	\end{subfigure}
	\caption{}
	\label{fig:learningcurves}

@@

(AL,RUD)=(1,1) Train R$^{2}$ 0.62/test R$^{2}$ 0.62
(AL,RUD)=(1,3) Train R$^{2}$ 0.85/test R$^{2}$ 0.85
(AL,RUD)=(1,5) Train R$^{2}$ 0.82/test R$^{2}$ 0.83
(AL,RUD)=(3,1) Train R$^{2}$ 0.71/test R$^{2}$ 0.71
(AL,RUD)=(5,1) Train R$^{2}$ 0.77/test R$^{2}$ 0.77

Talk about the cost of going wider, and inherently deeper

@@

>>>>

Next, we use learning curves, where the number of training points is varied, while the number of training points is varied to explore claims that PQCs offer better model performance using fewer training points and hyperparameters.\cite{schuld_circuit-centric_2020,du_expressive_2020,suzuki_predicting_2020}
For the BSE49 dataset, using 5 qubits, we found that M-M-CZ{\_}HWE-CNOT offers less overfit models, but offer poorer performance regarding the training and test set R$^{2}$s, as highlighted in Fig. \ref{fig:BSE5_learning_curves}.
Across the learning curve, the \textit{k}-NN model has a mean R$^{2}$s of 0.9998 for the training set and 0.5822 test set, while M-M-CZ{\_}HWE-CNOT has a mean R$^{2}$ of 0.1405 for the training set and R$^{2}$ of 0.136 for the test set.
For M-M-CZ{\_}HWE-CNOT, the training set performance peaks when using 10\% of the training data with an R$^{2}$ of 0.2015, dips to an R$^{2}$ of -0.0172 using 30\% of the training data, and then maintains an R$^{2}$ less than 0.2 for the remaining percentages of the training data.
Unlike the training points, the test reaches peak performance when 70\% of the total data is used in the training set, with a test R$^{2}$ of 0.2064.

When we analyze the effects of the number of training points on the DDCC dataset using the A2{\_}HWE-CNOT PQC, we found that

A2{\_}HWE-CNOT has $3nvk$ trainable parameters

Minimum number of classical parameters is $n+1$

\begin{figure}[H]
	\centering	
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/BSE/fivequbit/BSE_learningcurve}
		\caption{}
		\label{fig:BSE5_learning_curves}
	\end{subfigure}	
	\hfill
	\begin{subfigure}{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/DDCC/DDCC_learning_curves}
		\caption{}
		\label{fig:ddcclearningcurves}
	\end{subfigure}
	\caption{Learning curves for the (a) BSE49 database using the M-M-CZ{\_}HWE-CNOT circuit and (b) the DDCC dataset using the A2{\_}HWE-CNOT circuit. The x-axis denotes the percentage of training points, while the test set is held to 20\% of the total data. The y-axis denotes the R$^{2}$ of the training and test learning curves. Inlay plots show the training and test data for the classical models examined, where the bottom error bar denotes the minimum value, line the mean value, and top error bar the maximum value.}
	\label{fig:learningcurves}

@@

@@

\begin{table}[h!]
	\centering
	\small
	\begin{tabular}{|c|c|c|c|}
		\hline
		\textbf{Model} & \textbf{\# of Trainable Parameters} & \textbf{\# of Two-Qubit Gates} & \textbf{Circuit Depth} \\
		\hline
		A2\_HWE-CNOT & \(3nvk\) & \(nvk\) & \(vnk + 3vk + 2kl\) \\
		\hline
		M-M-CZ\_HWE-CNOT & \(3nvk\) & \(2nkl + nvk\) & \(2nkl + 2kl + nvk + 3vk\) \\
		\hline
		Linear, Ridge, Lasso, ElasticNet & \(n + 1\) & --- & --- \\
		\hline
	\end{tabular}
	\caption{Comparison of models with respect to parameter count, two-qubit gate count, and circuit depth.}
\end{table}
